---
title: "p8105_hw2_yc4016"
output: github_document
date: "2022-09-27"
author: "Yishan Chen"
---

```{r, setup, include = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1 

```{r}
hw2p1_df = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```


##### Description 

These data are not "tidy", becasue the route number is not a single variable. So we need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 


```{r}
hw2p1_df %>% 
  select(station_name, line) %>% 
  distinct
```

The ode chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.


```{r}
hw2p1_df %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Filter according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 


```{r}
hw2p1_df %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
hw2p1_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

hw2p1_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

##### Description 

* There are 465 distinct stations
* There are 84 stations are ADA compliant
* The proportion of station entrances / exits without vending allow entrance is 0.37
* There are 60 distinct stations serve the A train
* Of the stations that serve the A train, there are 17 ADA compliant.




# Problem 2 

### Read in and clean dataset mr_trash_wheel

```{r, mr_trash_wheel}
mr_trash_wheel= 
  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = as.integer(sports_balls),
    dataset = "mr_trash_wheel",
    year = as.numeric(year))%>% 
    relocate(dataset)
```

##### Description

* The size of the dataset are `r nrow(mr_trash_wheel)` rows and `r ncol(mr_trash_wheel)` columns.
* The name of variables are `r names(mr_trash_wheel)`. 


### Read in and clean dataset professor_trash_wheel

```{r, professor_trash_wheel}
professor_trash_wheel=
  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M96") %>% 
  janitor::clean_names() %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  drop_na(dumpster) %>% 
  mutate(dataset = "professor_trash_wheel")%>% 
  relocate(dataset)
```

##### Description

* The size of the dataset are `r nrow(professor_trash_wheel)` rows and `r ncol(professor_trash_wheel)` columns.
* The name of variables are `r names(professor_trash_wheel)`. 


### Combine datasets Mr_Trash_Wheel and Professor_Trash_Wheel

```{r, Combine datasets Mr_Trash_Wheel and Professor_Trash_Wheel}
combine=
  bind_rows(mr_trash_wheel, professor_trash_wheel)
```

```{r}
mr_trash_wheel_2020 =
  mr_trash_wheel %>% 
  filter(year == 2020)
```


##### Description 

* The size of the dataset are `r nrow(combine)` rows and `r ncol(combine)` columns.
* There are 641 observations. 
* The name of variables are `r names(combine)`. 
* The total weight of trash collected by Professor Trash Whee is `r sum(pull(professor_trash_wheel,weight))`
* The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(pull(mr_trash_wheel_2020, sports_balls))`


# Problem 3

### Read in and clean dataset pols-month.

```{r, pols_month}
pols_month = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-" ) %>% 
  mutate(
    year = as.numeric(year),
    month= month.name[as.numeric(month)],
    president = ifelse(prez_dem == 1, "demoncratic", "republican")) %>% 
  select(-day, -prez_dem, -prez_gop)
```

### Read in and clean dataset snp.

```{r, snp}
snp = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-")%>% 
  select(-day) %>% 
  mutate(
    year = as.numeric(year),
    month = month.name[as.numeric(month)] ) %>% 
  relocate(year, month)
```


